---
title: "project1"
output:
    html_document:
        keep_md: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

The following code is used to preprocess and prepare the data.

```{r echo=T}
library(ggplot2)
download.file(url = 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip', 
              destfile = paste(getwd(), 'activity.zip', sep = '/'))
unzip('activity.zip')
activity = data.table::fread('activity.csv')
```

### Histogram of the total number of steps taken each day

Following is the code to generate the histogram with the plot itself as well.

```{r, echo=TRUE}
total_steps = activity[, lapply(.SD, sum, na.rm=F), .SDcols = c('steps'), by = date] 
ggplot(total_steps) +
    geom_histogram(aes(x=steps), binwidth = 1000) +
    labs(title='Num of steps taken each day') + 
    labs(x='Steps', y= 'Frequency')

```

## What is mean total number of steps taken per day?
Following is the code to calculate the mean and median of the steps taken each day.

```{r, echo = TRUE}
total_steps[, .(Mean_Steps = mean(steps, na.rm = TRUE), Median_Steps = median(steps, na.rm = TRUE))]
```

### Time series plot of the average number of steps taken
Following is the code for the time series plot of steps taken each day.

```{r, echo = TRUE}

time_series = activity[, lapply(.SD, mean, na.rm=T), .SDcols = c('steps'), by = 'interval']

ggplot(data = time_series) + 
    geom_line(aes(x = interval, y = steps)) + 
    xlab("")

```

## What is the average daily activity pattern?
Following is the code for determining which interval has the max number of steps.
```{r,echo=T}
time_series[which.max(time_series$steps)]
```

## Imputing missing values
Here I will replace the missing values with the mean.

```{r, echo=TRUE}
activity[is.na(steps), 'steps'] = activity[, lapply(.SD, mean, na.rm=T), .SDcols=c('steps')]

```

### Histogram of the total number of steps taken each day after missing values are imputed
Here is the histogram of the steps after replacing missing values.
```{r, echo=TRUE}
total_steps = activity[, lapply(.SD, sum, na.rm=F), .SDcols = c('steps'), by = date] 
ggplot(total_steps) +
    geom_histogram(aes(x=steps), binwidth = 1000)+
    labs(title='Num of steps taken each day') + 
    labs(x='Steps', y= 'Frequency')

```


## Are there differences in activity patterns between weekdays and weekends?
Now lets compare the activity across weekdays and weekends.

```{r, echo = T}
#Reading a clean version of the data
activity = data.table::fread('activity.csv')
activity[, isWeekend := chron::is.weekend(activity$date)]
activity[is.na(steps), 'steps'] = activity[, lapply(.SD, mean, na.rm=T), .SDcols = c('steps')]
interval = activity[,lapply(.SD, mean, na.rm=T), .SDcols=c('steps'), by=.(interval, isWeekend)]
```
Now to create the plot
```{r echo=T}
ggplot(data=interval) +
    geom_line(aes(x=interval,y=steps,color=isWeekend)) +
    labs(x='Interval', y='Steps') +
    labs(main='Comparing activity over weekends and weekdays') +
    facet_wrap(~'isWeekend', ncol=1, nrow=2)
```



